Factors Influencing Early Childhood Comprehension of American English with a Bayesian Model

### 2024-12-03 paper/paper.qmd
update paper

Early childhood language comprehension is a fundamental area of cognitive development that influences future literacy and communication skills. Research has identified several key predictors of language comprehension in young children. For example, a study by Taylor et al.(@Zubrick_Taylor_Christensen_2015) found that socioeconomic factors, family background, and personal characteristics significantly influence early childhood receptive vocabulary development. As highlighted by Lawton et al.(@Lawton5129), the quality and quantity of language exposure in infancy are associated with brain development and long-term language achievement. In addition, Bergelson's study(@Zonarich_2024) showed that babies begin to understand common nouns as early as 6 to 7 months of age, which suggests that comprehension begins earlier than previously thought. Despite these observations, we believe there is still a need for robust research that considers the range of demographic, developmental, and environmental factors that influence the early comprehension of American English learners.

@@

My modelling approach aims to quantify the relationship between various child and caregiver characteristics and comprehension in early childhood. For this analysis, we use a Bayesian linear model to examine how factors such as age, production (number of words produced) influence comprehension. The model is implemented using the `stan glm` function, with a Gaussian distribution to capture the variability in comprehension.

@@

The Communicative Development Inventory (CDI) is a standardized assessment tool used to study early language development in children aged 8 to 30 months. Originally developed by @citewordbankproject, the CDI provides a reliable and general method for evaluating both vocabulary comprehension and production based on parent reports. This chapter revisits the strengths and limitations of CDI, its psychometric properties, and the reliability of the instrument in longitudinal studies.

>>>>

Early childhood language comprehension is a fundamental area of cognitive development that influences future literacy and communication skills. Research has identified several key predictors of language comprehension in young children. For example, a study by Taylor,  found that socioeconomic factors, family background, and personal characteristics significantly influence early childhood receptive vocabulary development [@Zubrick_Taylor_Christensen_2015]. As highlighted by Lawton, the quality and quantity of language exposure in infancy are associated with brain development and long-term language achievement[@Lawton5129]. In addition, Bergelson's study showed that babies begin to understand common nouns as early as 6 to 7 months of age, which suggests that comprehension begins earlier than previously thought[@Zonarich_2024]. Despite these observations, we believe there is still a need for robust research that considers the range of demographic, developmental, and environmental factors that influence the early comprehension of American English learners.

@@

My modelling approach aims to quantify the relationship between various child and caregiver characteristics and comprehension in early childhood. For this analysis, we use a Bayesian linear model to examine how factors such as age, production (number of words produced) influence comprehension. The model is implemented using the `stan_glm` function, with a Gaussian distribution to capture the variability in comprehension.

@@

The Communicative Development Inventory (CDI) is a standardized assessment tool used to study early language development in children aged 8 to 30 months. Originally developed by [@citewordbankproject], the CDI provides a reliable and general method for evaluating both vocabulary comprehension and production based on parent reports. This chapter revisits the strengths and limitations of CDI, its psychometric properties, and the reliability of the instrument in longitudinal studies.
